Modularly Composing Typed Language Fragments (Supplemental Material)

### 2014-11-04 att-pldi15/att-pldi15-supplement.tex
Updated TR, related work, cleaned up obdy

%\documentclass[12pt]{article}
\documentclass[10pt,preprint]{sigplanconf}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{mathpartir}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

@@

\conferenceinfo{-}{-} 
\copyrightyear{-} 
\copyrightdata{[to be supplied]}

@@

\title{Modularly Composing Typed Language Fragments}
\subtitle{Supplemental Material}
\authorinfo{}{}{}
%\authorinfo{Cyrus Omar \and Jonathan Aldrich}
 %         {Carnegie Mellon University}
  %         {\{comar, aldrich\}@cs.cmu.edu}

@@

We use {$\mathcal{L}\{{\rightharpoonup}\,{\forall}\,{\mu}\,{\iunit}\,{\times}\,{+}\}$}, the syntax for which is shown in Figure \ref{syntax-IL}, as representative of any intermediate language for a typed functional language. In fact, our intention is not to prescribe a particular choice of IL, so we will here only review the key metatheoretic properties that the IL must possess. Each choice of IL is technically a distinct dialect of @$\lambda$, but for the broad class of ILs that enjoy these properties, the metatheory in the remainder of the supplement should follow without trouble.

@@

\begin{lemma}[Internal Progress]
If $\emptyset~\emptyset \vdash \iota : \tau$ then either $\iota~\texttt{val}$ or $\iota \mapsto \iota'$.
\end{theorem}
\begin{lemma}[Internal Preservation]
If $\emptyset~\emptyset \vdash \iota : \tau$ and $\iota \mapsto \iota'$ then $\emptyset~\emptyset \vdash \iota' : \tau$. 
\end{lemma}

@@

\begin{figure}[t]
\small
\hspace{-5px}\input{syntax-table-IL}
\caption{Syntax of {$\mathcal{L}\{{\rightharpoonup}\,{\forall}\,{\mu}\,{\iunit}\,{\times}\,{+}\}$}, our internal language (IL). Metavariable $x$ ranges over term variables and $\alpha$ and $t$ both range over type variables.}
\label{syntax-IL}
\end{figure}

\begin{figure}
\small\fbox{$\Delta \vdash \delta : \Delta$}~~$\delta ::=  \emptyset ~|~ \delta, \tau/\alpha$

\begin{mathpar}
\inferrule[tysub-emp]{ }{\Delta \vdash \emptyset : \emptyset}

\inferrule[tysub-ext]{
    \Delta \vdash \delta : \Delta'\\
    \Delta \vdash \tau
}{
    \Delta \vdash \delta, \tau/\alpha : \Delta', \alpha
}
\end{mathpar}
\caption{Internal Type Substitution Validity}
\label{fig:internal-type-substitution-validity}
\end{figure}
\begin{figure}
\small\fbox{$\Delta~\Gamma \vdash \gamma : \Gamma$}~~$\gamma ::= \emptyset ~|~ \gamma, \iota/x$
\begin{mathpar}
\inferrule[tmsub-emp]{ }{\Delta~\Gamma \vdash \emptyset : \emptyset}

\inferrule[tmsub-ext]{
    \Delta~\Gamma \vdash \gamma : \Gamma'\\
    \Delta \vdash \tau\\
    \Delta~\Gamma \vdash \iota : \tau
}{
    \Delta~\Gamma \vdash \gamma, \iota/x : \Gamma', x : \tau
}
\end{mathpar}
\caption{Internal Term Substitution Validity}
\label{fig:internal-term-substitution-validity}
\end{figure}
\begin{lemma}[Internal Type Substitution on Types] If $\Delta \vdash \delta : \Delta'$ and $\Delta\Delta' \vdash \tau$ then $\Delta \vdash [\delta]\tau$.\end{lemma}

@@

@@

\section{Tycon Contexts}
\begin{figure}\fbox{$\vdash \Phi$}\vspace{-25px}
\begin{mathpar}
\small
\inferrule[tcc-emp]{ }{\vdash \cdot}

\inferrule[tcc-ext]{
    \vdash \Phi\\
    \tc \notin \text{dom}(\Phi)\\
    \keq{\emptyset}{\ktyidx}\\
    \sofkz{\emptyset}{\emptyset}{\Phi}{\stx{schema}}{\ktyidx \rightarrow \kity}\\\\
    \vdash_{\Phi,  \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}} \omega \sim \tcsig{\ktyidx}{\chi}
}{
    \vdash \Phi, \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}
}
\end{mathpar}
\caption{Tycon context well-definedness.}
\label{fig:tycon-ctxs}
\end{figure}
\begin{figure}[t]
\small\fbox{$\vdash_\Phi \omega \sim \psi$}
\begin{mathpar}\small
\inferrule[ocstruct-intro]{
    \introsig{\klitidx} \in \chi\\
    \emptyset \vdash \klitidx\\\\
    \sofkz{\emptyset}{\emptyset}{\Phi}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow \kitm}
}{
    \vdash_\Phi \tcstructn{\stx{def}} \sim \tcsig{\ktyidx}{\chi}
}

\inferrule[ocstruct-targ]{
    \vdash_\Phi \omega \sim \tcsig{\ktyidx}{\chi}\\
    \opname{op}\notin\text{dom}(\chi)\\
    \emptyset \vdash \klitidx\\\\
    \sofkn{\emptyset}{\emptyset}{\Phi}{0}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow (\kty \times \kitm)}
}{
    \vdash_\Phi \tcstructc{\omega}{op}{\stx{def}} \sim \tcsig{\ktyidx}{\chi, \opsig{\opname{op}}{\klitidx}}
}
\end{mathpar}
\caption{Checking opcon structures against tycon signatures.}
\label{ocstruct}
\end{figure}

\subsection{Tycon Context Well-Definedness}

@@

\subsection{Full Examples}

\subsection{Kinding Context Formation}

@@

\subsection{Dynamic Semantics}

\subsection{Kind Safety}

\section{Types}
\subsection{Type Translations}

@@

\paragraph{Unicity}
The rules are structured so that if a term is well-typed, both its type and translation are unique.\todo{could move this whole thing to supplement if room needed}
\begin{theorem}[Unicity]
If $\vdash \Phi$ and $\vdash_\Phi \Upsilon \leadsto \Gamma$ and $\vdash_\Phi \st \leadsto \tau$ and $\vdash_\Phi \st' \leadsto \tau'$ and $\eanaX{e}{\st}{\iota}$ and $\eanaX{e}{\st'}{\iota'}$ then $\st = \st'$ and $\tau=\tau'$ and $\iota = \iota'$.
\end{theorem}

\section{External Language}
\subsection{Additional Desugarings}
\subsection{Typing}
\subsection{Proof of Regular String Soundness Tycon Invariant}

@@

\inferrule[args-s]{
    \keyw{args}(\es)=_{n} \st
}{
    \keyw{args}(\es;e)=_{n+1} \skap{\karg}{\svar{rcons}}~\st~(\slam{\kty}{\svar{ty}}{\sana{n}{\svar{ty}}},\slam{\kunit}{\_}{\ssyn{n}})
}
\end{mathpar}
We assume that the definitions of the standard helper functions $\small\svar{nil} :: \kforall{\kalpha}{\klist{\kalpha}}$ and $\small\svar{rcons} :: \kforall{\kalpha}{\klist{\kalpha} \rightarrow \kalpha \rightarrow \klist{\kalpha}}$, which adds an item to the end of a list, have been substituted into these rules. The result is that the $n$th element of the argument interface list simply wraps the static terms $\sana{n}{\st}$ and $\ssyn{n}$.\end{document}

>>>>

%\documentclass[12pt]{article}
\documentclass[12pt]{article}
\usepackage{fullpage,cmu-titlepage2}
\usepackage{times}
% The following \documentclass options may be useful:
%
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% authoryear    To obtain author/year citation style instead of numeric.
\usepackage{amsmath}
\usepackage{amssymb} 
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{mathpartir}
\usepackage{enumerate}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

@@

@@

\title{Modularly Composing Typed Language Fragments (Supplemental Material)}
%\subtitle{Supplemental Material}
%\authorinfo{}{}{}
%\authorinfo{Cyrus Omar \and Jonathan Aldrich}
 %         {Carnegie Mellon University}
  %         {\{comar, aldrich\}@cs.cmu.edu}

\author{ }

\abstract{
    This document provides the full technical development described in the paper ``Modularly Composing Typed Language Fragments'' submitted to PLDI 2015. 
}
\keywords{Statistical Model Checking, Markov Decision Processes, Reinforcement Learning}

\trnumber{ }

@@

We use {$\mathcal{L}\{{\rightharpoonup}\,{\forall}\,{\mu}\,{\iunit}\,{\times}\,{+}\}$}, the syntax for which is shown in Figure \ref{syntax-IL}, as representative of any intermediate language for a typed functional language.

\begin{figure}[h!]
\small
\hspace{-5px}\input{syntax-table-IL}
\caption{Syntax of {$\mathcal{L}\{{\rightharpoonup}\,{\forall}\,{\mu}\,{\iunit}\,{\times}\,{+}\}$}, our internal language (IL). Metavariable $x$ ranges over term variables and $\alpha$ and $t$ both range over type variables.}
\label{syntax-IL}
\end{figure}

In fact, our intention is not to prescribe a particular choice of IL, so we will here only review the key metatheoretic properties that the IL must possess. Each choice of IL is technically a distinct dialect of @$\lambda$, but for the broad class of ILs that enjoy these properties, the metatheory in the remainder of the supplement should follow without trouble.

@@

\begin{lemma}[Internal Progress]
If $\emptyset~\emptyset \vdash \iota : \tau$ then either $\iota~\texttt{val}$ or $\iota \mapsto \iota'$.
\end{lemma}
\begin{lemma}[Internal Preservation]
If $\emptyset~\emptyset \vdash \iota : \tau$ and $\iota \mapsto \iota'$ then $\emptyset~\emptyset \vdash \iota' : \tau$. 
\end{lemma}

@@

\begin{figure}[h]
\small\fbox{$\Delta \vdash \delta : \Delta$}~~$\delta ::=  \emptyset ~|~ \delta, \tau/\alpha$
\begin{mathpar}
\inferrule[tysub-emp]{ }{\Delta \vdash \emptyset : \emptyset}

\inferrule[tysub-ext]{
    \Delta \vdash \delta : \Delta'\\
    \Delta \vdash \tau
}{
    \Delta \vdash \delta, \tau/\alpha : \Delta', \alpha
}
\end{mathpar}
\caption{Internal Type Substitution Validity}
\label{fig:internal-type-substitution-validity}
\end{figure}

\begin{lemma}[Internal Type Substitution on Types] If $\Delta \vdash \delta : \Delta'$ and $\Delta\Delta' \vdash \tau$ then $\Delta \vdash [\delta]\tau$.\end{lemma}

@@

\begin{figure}[h]
\small\fbox{$\Delta~\Gamma \vdash \gamma : \Gamma$}~~$\gamma ::= \emptyset ~|~ \gamma, \iota/x$
\begin{mathpar}
\inferrule[tmsub-emp]{ }{\Delta~\Gamma \vdash \emptyset : \emptyset}

\inferrule[tmsub-ext]{
    \Delta~\Gamma \vdash \gamma : \Gamma'\\
    \Delta \vdash \tau\\
    \Delta~\Gamma \vdash \iota : \tau
}{
    \Delta~\Gamma \vdash \gamma, \iota/x : \Gamma', x : \tau
}
\end{mathpar}
\caption{Internal Term Substitution Validity}
\label{fig:internal-term-substitution-validity}
\end{figure}

@@

\section{Tycons}
\begin{figure}[h]
\small
$\begin{array}{lrcl}
\textbf{tycons} & c & ::= & \rightharpoonup ~|~ \tc ~|~ \keyw{other}[m]\\
\textbf{tycon contexts} & \Phi & ::= & \cdot ~|~ \Phi, \tcdef{\tc}{\psi}{\theta}\\
\textbf{tycon structures} & \theta & ::= & \tcstruct{\st}{\omega} \\
\textbf{opcon structures} & \omega & ::= & \tcstructn{\st} ~|~ \tcstructc{\omega}{op}{\st}\\
\textbf{tycon sigs} & \psi & ::= & \tcsig{\kappa}{\chi}\\
\textbf{opcon sigs}& \chi & ::= & \introsig{\kappa} ~|~ \opsigS{\chi}{op}{\kappa}\\
\end{array}$
\caption{Syntax of tycons. Metavariables $\tc$ and $\opname{op}$ range over user-defined tycon and opcon names, respectively, and $m$ ranges over natural numbers.}
\label{syntax-TC}
\end{figure}

\subsection{Tycon Contexts}
Tycon contexts are ordered mappings from tycon names $\tc$ to tycon definitions. The tycon context well-definedness judgement $\vdash \Phi$ is specified in Figure \ref{fig:tycon-ctxs}.

\begin{figure}[h]
\fbox{$\vdash \Phi$}\vspace{-25px}
\begin{mathpar}
\small
\inferrule[tcc-emp]{ }{\vdash \cdot}

\inferrule[tcc-ext]{
    \vdash \Phi\\
    \tc \notin \text{dom}(\Phi)\\
    \keq{\emptyset}{\ktyidx}\\
    \sofkz{\emptyset}{\emptyset}{\Phi}{\stx{schema}}{\ktyidx \rightarrow \kity}\\\\
    \vdash_{\Phi,  \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}} \omega \sim \tcsig{\ktyidx}{\chi}
}{
    \vdash \Phi, \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}
}
\end{mathpar}
\caption{Tycon context well-definedness.}
\label{fig:tycon-ctxs}
\end{figure}

Opcon structures are checked against the tycon's signature.

\begin{figure}[h]
\small\fbox{$\vdash_\Phi \omega \sim \psi$}
\begin{mathpar}\small
\inferrule[ocstruct-intro]{
    \introsig{\klitidx} \in \chi\\
    \emptyset \vdash \klitidx\\\\
    \sofkz{\emptyset}{\emptyset}{\Phi}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow \kitm}
}{
    \vdash_\Phi \tcstructn{\stx{def}} \sim \tcsig{\ktyidx}{\chi}
}

\inferrule[ocstruct-targ]{
    \vdash_\Phi \omega \sim \tcsig{\ktyidx}{\chi}\\
    \opname{op}\notin\text{dom}(\chi)\\
    \emptyset \vdash \klitidx\\\\
    \sofkn{\emptyset}{\emptyset}{\Phi}{0}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow (\kty \times \kitm)}
}{
    \vdash_\Phi \tcstructc{\omega}{op}{\stx{def}} \sim \tcsig{\ktyidx}{\chi; \opsig{\opname{op}}{\klitidx}}
}
\end{mathpar}
\caption{Checking opcon structures against tycon signatures.}
\label{ocstruct}
\end{figure}

For the purposes of the SL, only the type signatures are relevant. We thus define  judgements $\vdash \psi$, $\vdash \chi$ and $\vdash \Phi~\mathtt{sigsok}$ in Figure \ref{fig:tycon-and-opcon-signature-well-formedness} for checking this only.

The following lemmas characterize these judgements. 
\begin{figure}[h]
\fbox{$\vdash \psi$}\vspace{-25px}
\begin{mathpar}
\small
\inferrule[tcsig-ok]{
    \keq{\emptyset}{\ktyidx}\\
    \vdash \chi
}{
    \vdash \tcsig{\ktyidx}{\chi}
}
\end{mathpar}
\fbox{$\vdash \chi$}\vspace{-25px}
\begin{mathpar}
\small
\inferrule[ocsig-intro-ok]{
    \emptyset \vdash \kappa
}{
    \vdash \introsig{\kappa}
}

\inferrule[ocsig-targ-ok]{
    \emptyset \vdash \kappa\\
    \opname{op} \notin \text{dom}(\chi)
}{
    \vdash \opsigS{\chi}{op}{\kappa}
}
\end{mathpar}
\fbox{$\vdash \Phi~\mathtt{sigsok}$}\vspace{-25px}
\begin{mathpar}
\small
\inferrule[tcc-sigsok-emp]{ }{
    \vdash \emptyset~\mathtt{sigsok}
}

\inferrule[tcc-sigsok-ext]{
    \vdash \Phi~\mathtt{sigsok}\\
    \vdash \psi
}{
    \vdash \Phi, \tcdef{\tc}{\psi}{\theta}~\mathtt{sigsok}
}
\end{mathpar}
\caption{Tycon and Opcon Signature Well-Formedness}
\label{fig:tycon-and-opcon-signature-well-formedness}
\end{figure}

\begin{lemma}[Inversion on Tycon Context Well-Definedness]
\label{lemma:inversion-on-tycon-context-well-definedness}
If (1) $$\vdash \Phi, \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}$$ then (a) $\vdash \Phi$ and (b) $\tc \notin \text{dom}(\Phi)$ and (c) $\keq{\emptyset}{\ktyidx}$ and (d) $\sofkz{\emptyset}{\emptyset}{\Phi}{\stx{schema}}{\ktyidx \rightarrow \kity}$ and (e) $\vdash_{\Phi,  \tcdef{\tc}{\tcsig{\ktyidx}{\chi}}{\tcstruct{\stx{schema}}{\omega}}} \omega \sim \tcsig{\ktyidx}{\chi}$.
\end{lemma}
\begin{proof}
Rule induction on assumption (1). Rule (tcc-emp) does not apply. Rule (tcc-ext) applies. The conclusions (a-e) are the five premises of (tcc-ext) and thus follow immediately.
\end{proof}

\begin{lemma}[Tycon Context Composition]
If (i) $\vdash \Phi_1$ and (ii) $\vdash \Phi_2$ and (iii) $\text{dom}(\Phi_1) \cap \text{dom}(\Phi_2) = \emptyset$ then (a) $\vdash \Phi_1\Phi_2$.
\end{lemma}
\begin{proof}
By syntactic case analysis on $\Phi_1$ and $\Phi_2$.

If $\Phi_1=\emptyset$ and $\Phi_2=\emptyset$, then $\Phi_1\Phi_2=\emptyset$ and (a) follows by (tcc-emp).

If $\Phi_2=\emptyset$ then $\Phi_1\Phi_2=\Phi_1$ and (a) follows by (1).

If $\Phi_1=\emptyset$ then $\Phi_1\Phi_2=\Phi_2$ and (a) follows by (2).

If $\Phi_1=\Phi_1', \tcdef{\tc_1}{\psi_1}{\theta_1}$ and $\Phi_2=\Phi_2',\tcdef{\tc_2}{\psi_2}{\theta_2}$ then $\Phi_1\Phi_2=\Phi_1'\Phi_2',\tcdef{\tc_1}{\psi_1}{\theta_1},\tcdef{\tc_2}{\psi_2}{\theta_2}$. Let (1-5) be the result of applying Lemma \ref{lemma:inversion-on-tycon-context-well-definedness} on (i), and (6-10) be the result of applying Lemma \ref{lemma:inversion-on-tycon-context-well-definedness} on (ii). By (iii) and the usual properties of domains of finite mappings, we have that (11) $\text{dom}(\Phi_1') \cap \text{dom}(\Phi_2') = \emptyset$ and further by (2) we have that (12) $\tc\notin\keyw{dom}(\Phi_1'\Phi_2')$ and by (8) that (13) $\tc'\notin\text{dom}({\Phi_1'\Phi_2'})$.  By the IH on (1), (6) and (11) we have that (14) $\vdash \Phi_1'\Phi_2'$. By (tcc-ext) on (14), (12), (3), (4) and \todo{weakening of kinding and opcon well-definedness} on (5), we have that (15) $\vdash \Phi_1'\Phi_2', \tcdef{\tc_1}{\psi_1}{\theta_1}$. By (tcc-ext) on (15), (13), (8), (9) and SAME on (10), we have (a).
\end{proof}

\begin{lemma}[Intro Opcon Existence and Well-Definedness]
If (1) $\vdash_\Phi \omega \sim \tcsig{\ktyidx}{\chi}$ then (a) $\keyw{intro}[\klitidx] \in \chi$ and (b) $\emptyset \vdash \klitidx$ and (c) $\keyw{ana~intro}=\stx{def}\in\omega$ and (d) $\sofkz{\emptyset}{\emptyset}{\Phi}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow \kitm}$.
\end{lemma}
\begin{proof}
Rule induction on assumption (1). If rule (ocstruct-intro) applies then conclusion (c) follows syntactically and (a), (b) and (d) are the three premises and thus follow immediately. If rule (ocstruct-targ) applies, then we apply the IH to the first premise.
\end{proof}

\begin{lemma}[Targeted Opcon Well-Definedness and Unicity]
If (1) $\vdash_\Phi \omega \sim \tcsig{\ktyidx}{\chi}$ and (2) $\keyw{syn}~\opname{op}=\stx{def} \in \omega$ then (a) $\opname{op}[\klitidx] \in \chi$ and (b) $\emptyset \vdash \klitidx$ and (c) $
    \sofkn{\emptyset}{\emptyset}{\Phi}{0}{\stx{def}}{\ktyidx \rightarrow \klitidx \rightarrow \kargs \rightarrow (\kty \times \kitm)}$ and (d) if $\keyw{syn}~\opname{op}=\stx{def}' \in \omega$ then $\stx{def}'=\stx{def}$ and (e) if $\opname{op}[\klitidx']\in\chi$ then $\klitidx'=\klitidx$.
\end{lemma}
\begin{proof}
Rule induction on assumption (1). Rule (ocstruct-intro) does not apply by assumption (2). Rule (ocstruct-targ) applies. Conclusion (a) follows syntactically. Conclusions (b) and (c) are premises 3 and 4. Conclusion (d) follows because either $\stx{def}'=\stx{def}$ (i.e. we considering the current definition) or we can apply the IH to premise 1 and the assumption of (d). Conclusion (e) follows because premise 2 checks the unicity condition directly.
\end{proof}

\subsection{Full Examples}
\subsubsection{Regular Strings}

\subsubsection{Labeled Products}

\subsubsection{Records}

\begin{figure}[h]
\small\fbox{$\kDelta \vdash \kappa$}
\begin{mathpar}
\inferrule[kf-arrow]{
    \kDelta \vdash \kappa_1\\
    \kDelta \vdash \kappa_2
}{
    \kDelta \vdash \karrow{\kappa_1}{\kappa_2}
}

\inferrule[kf-alpha]{
    \kalpha \in \kDelta
}{
    \kDelta \vdash \kalpha
}

\inferrule[kf-forall]{
    \kDelta, \kalpha \vdash \kappa
}{
    \kDelta \vdash \kforall{\kalpha}{\kappa}
}

\inferrule[kf-k]{
    \kvar \in \kDelta
}{
    \kDelta \vdash \kvar
}

\inferrule[kf-ind]{
    ???
}{
    \kDelta \vdash \kmu{\kvar}{\kappa}
}

\inferrule[kf-unit]{ }{
    \kDelta \vdash \kunit
}

\inferrule[kf-prod]{
    \kDelta \vdash \kappa_1\\
    \kDelta \vdash \kappa_2
}{
    \kDelta \vdash \kprod{\kappa_1}{\kappa_2}
}

\inferrule[kf-sum]{\kDelta \vdash \kappa_1\\
    \kDelta \vdash \kappa_2
}{
    \kDelta \vdash \ksum{\kappa_1}{\kappa_2}
}

\inferrule[kf-ty]{ }{\kDelta \vdash \kty}

\inferrule[kf-ity]{ }{\kDelta \vdash \kity}

\inferrule[kf-itm]{ }{\kDelta \vdash \kitm}
\end{mathpar}
\caption{Kind Formation}
\label{fig:kind-formation}
\end{figure}
\begin{lemma}[Kind Variable Substitution - Kinds]~\\
\begin{enumerate}
\item If $\kDelta, \kalpha \vdash \kappa$ and $\kDelta \vdash \kappa'$ then $\kDelta \vdash [\kappa'/\kalpha]\kappa$.
\item If $\kDelta, \kvar \vdash \kappa$ and $\kDelta \vdash \kappa'$ then $\kDelta \vdash [\kappa'/\kvar]\kappa$.
\end{enumerate}
\end{lemma}
\begin{figure}[h]
???
\caption{Positive Kinds}
\label{fig:positive-kinds}
\end{figure}

\begin{figure}[h]
\small\fbox{$\kDelta \vdash \kGamma$}
\begin{mathpar}
\inferrule[kctx-emp]{ }{\kDelta \vdash \emptyset}

\inferrule[kctx-ext]{
    \kDelta \vdash \kGamma\\
    \kDelta \vdash \kappa
}{
    \kDelta \vdash \kGamma, \sx :: \kappa
}
\end{mathpar}
\caption{Kinding Context Formation}
\label{fig:kinding-context-formation}
\end{figure}

@@

\begin{lemma}[Equality Kind Well-Formedness]
If $\keq{\emptyset}{\kappa}$ then $\emptyset \vdash \kappa$. 
\end{lemma}

@@

\subsection{Static Dynamics}

\begin{lemma}[Static Canonical Forms]
...
\end{lemma}

Discussion of decidability and weak normalization.

\subsection{Type Translations}

@@

\subsection{Arguments}
\begin{definition}[Argument Interface Kind]
The kind abbreviated $\karg$ is defined as\[\karg := (\kty \rightarrow \kitm) \times (\kunit \rightarrow \kty\times\kitm)\]
\end{definition}

\subsection{Kind Safety}
\paragraph{Kind Safety} Kind safety ensures that normalization of well-kinded static terms cannot go wrong. We can take a standard progress and preservation based approach. 
\begin{theorem}[Static Progress]\label{thm:static-progress}
If $\sofkn{\emptyset}{\emptyset}{\Phi}{n}{\st}{\kappa}$ and $\vdash \Phi$ and $|\es|=n$ and $\vdash_\Phi \Upsilon \leadsto \Gamma$ then $\sval{\st}{\es; \Upsilon; \Phi}$ or $\serr{\st}{\es; \Upsilon; \Phi}$ or $\sstep{\st}{\es; \Upsilon; \Phi}{\st'}$.
\end{theorem}

\begin{theorem}[Static Preservation]\label{thm:static-preservation}
If $\sofkn{\emptyset}{\emptyset}{\Phi}{n}{\st}{\kappa}$ and $\vdash \Phi$ and $|\es|=n$ and $\vdash_\Phi \Upsilon \leadsto \Gamma$ and $\sstep{\st}{\es; \Upsilon; \Phi}{\st'}$ then $\sofkn{\emptyset}{\emptyset}{\Phi}{n}{\st'}{\kappa}$.
\end{theorem}
The case in the proof of Theorem \ref{thm:static-preservation} for $\ssyn{n}$ requires that the following theorem be mutually defined. The mutual induction is well-founded because the total number of argument lists in the terms being considered decreases.
\begin{theorem}[Type Synthesis]
If $\vdash \Phi$ and $\vdash_\Phi \Upsilon \leadsto \Gamma$ and $\esynX{e}{\st}{\iota}$ then $\vdash_\Phi \st \leadsto \tau$ (and thus $\istype{\st}{\Phi}$). 
\end{theorem}

\section{External Language}
\subsection{Additional Desugarings}
\subsection{Typing}
\paragraph{Unicity}
The rules are structured so that if a term is well-typed, both its type and translation are unique.\todo{could move this whole thing to supplement if room needed}
\begin{theorem}[Unicity]
If $\vdash \Phi$ and $\vdash_\Phi \Upsilon \leadsto \Gamma$ and $\vdash_\Phi \st \leadsto \tau$ and $\vdash_\Phi \st' \leadsto \tau'$ and $\eanaX{e}{\st}{\iota}$ and $\eanaX{e}{\st'}{\iota'}$ then $\st = \st'$ and $\tau=\tau'$ and $\iota = \iota'$.
\end{theorem}

\subsection{Proof of Regular String Soundness Tycon Invariant}

@@

\inferrule[args-s]{
    \keyw{args}(\es)=_{n} \st
}{
    \keyw{args}(\es;e)=_{n+1} \skap{\karg}{\svar{rcons}}~\st~(\slam{\kty}{\svar{ty}}{\sana{n}{\svar{ty}}},\slam{\kunit}{\_}{\ssyn{n}})
}
\end{mathpar}
We assume that the definitions of the standard helper functions $\small\svar{nil} :: \kforall{\kalpha}{\klist{\kalpha}}$ and $\small\svar{rcons} :: \kforall{\kalpha}{\klist{\kalpha} \rightarrow \kalpha \rightarrow \klist{\kalpha}}$, which adds an item to the end of a list, have been substituted into these rules. The result is that the $n$th element of the argument interface list simply wraps the static terms $\sana{n}{\st}$ and $\ssyn{n}$.

\begin{lemma}
\small If $\sofkn{\kDelta}{\kGamma}{\Phi}{n'}{\st}{\kappa}$ and $n > n'
$ then $\sofkn{\kDelta}{\kGamma}{\Phi}{n}{\st}{\kappa}$.
\end{lemma}

\end{document}
